Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-09-11 tex/scam/sparsistency.tex
minor edits

We divide our analysis into two parts. We first establish a sufficient deterministic condition for sparsistency.  We then consider the
stochastic setting and argue that the deterministic conditions hold with high probability. In all of our results and analysis, we let $c, C$ represent absolute constants; the actual values of $c,C$ could change from line to line.

@@

We construct an additive convex solution $\{\hat{d}_k\}_{k=1,...,p}$ that is zero for $k \in S^c$ and show that it satisfies the KKT conditions for optimality of optimization~\ref{opt:alternate_opt}. We define $\hat{d}_k$ for $k \in S$ to be a solution to the restricted regression (defined below). We will then also show that $\hat{c}_k = 0$ satisfies the optimality condition of optimization~\ref{opt:alternate_opt_concave} for all $k \in S^c$.

\begin{definition}
\label{def:restricted_regression}
We define as \emph{restricted regression} where we restrict the indices $k$ in
optimization (\ref{opt:alternate_opt}) to lie in the set $S$ instead of ranging from $1,...,p$:
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, ..., d_{k,n-1} \geq 0
\]
\end{definition}

\begin{theorem} (Deterministic setting)
\label{thm:deterministic}
Let $\{\hat{d}_k \}_{k \in S}$ be a minimizer of the restricted regression as defined above.
Let $\hat{r} \coloneqq Y - \sum_{k \in S} \bar{\Delta}_k \hat{d}_k$ be the restricted regression residual. \\

Suppose for all $k\in S^c$, for all $i=1,...,n$, $\lambda_n > | \frac{1}{2n}
\hat{r}^\tran \mathbf{1}_{(i:n)}|$ where $\mathbf{1}_{(i:n)}$ is 1 on the coordinates of the $i$-th largest to the $n$-th largest entries of $X_k$ and 0 elsewhere.\\

Then the following are true:
\begin{enumerate}
\item Let $\hat{d}_k = 0$ for $k \in S^c$, then \{$\hat{d}_k\}_{k=1,...,p}$ is an optimal solution to optimization~\ref{opt:alternate_opt}. Furthermore, any solution to the optimization program \ref{opt:alternate_opt} must be zero on $S^c$.
\item For all $k \in S^c$, the solution $\hat{c}_k$ to optimization~\ref{opt:alternate_opt_concave} must be 0 and must be unique.
\end{enumerate}

@@

This result holds regardless of whether we impose the boundedness conditions in optimization~\ref{opt:alternate_opt} and~\ref{opt:alternate_opt_concave} or not.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.\\

Theorem~\ref{thm:deterministic} allows us to analyze false negative
rates and false positive rates separately. To control false positives,
we analyze when the condition on $\lambda_n$ is fulfilled for all $j$ and all $k \in S^c$. To control false negatives, we analyze the restricted regression. \\

The proof of theorem~\ref{thm:deterministic} looks at KKT conditions of optimization~\ref{opt:alternate_opt}, similar to the now standard \emph{primal-dual witness} technique~\cite{wainwright2009sharp}. We cannot derive analogous \emph{mutual incoherence} conditions because the estimation is nonparametric -- even the low dimensional restricted regression has $s(n-1)$ variables. The details of the proof are in section~\ref{sec:deterministic_proof} of the Appendix.

@@

We let $S = \{ k = 1,...,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$.

@@

And therefore，for all $k \in S^c$, both the AC solution $\hat{f}_k$, from optimization~\ref{opt:alternate_opt}, and the DC solution $\hat{g}_k$, from optimization~\ref{opt:alternate_opt_concave} are zero. \\
\end{theorem}

@@

\begin{theorem} (Controlling false negatives) \\
\label{thm:false_negative}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution to the restricted regression with $B$-boundedness constraint and let $\hat{g}_k$'s be any DC solution to the restricted regression with $B$-boundedness constraint. Let $\tilde{\sigma}$ denote $\max(\sigma, B)$.\\

>>>>

We divide our analysis into two parts. We first establish a sufficient
deterministic condition for consistency of the sparsity pattern
selection procedure---a property sometimes called \textit{sparsistency}.
We then consider the
stochastic setting and argue that the deterministic conditions hold
with high probability. Note that in all of our results and analysis, we let $c,
C$ represent absolute constants; the actual values of $c,C$ may change from line to line.

@@

We construct an additive convex solution $\{\hat{d}_k\}_{k=1,...,p}$
that is zero for $k \in S^c$, where $S$ is the set of relevant
variables, and show that it satisfies the KKT
conditions for optimality of optimization~\eqref{opt:alternate_opt}. We
define $\hat{d}_k$ for $k \in S$ to be a solution to the restricted
regression (defined below). We also show that $\hat{c}_k =
0$ satisfies the optimality condition of
optimization~\eqref{opt:alternate_opt_concave} for all $k \in S^c$.

\begin{definition}
\label{def:restricted_regression}
We define the \emph{restricted regression} problem 
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, ..., d_{k,n-1} \geq 0
\]
where we restrict the indices $k$ in
optimization \eqref{opt:alternate_opt} to lie in the set $S$ of true
relevant variables.
\end{definition}

\begin{theorem}[Deterministic setting]
\label{thm:deterministic}
Let $\{\hat{d}_k \}_{k \in S}$ be a minimizer of the restricted regression as defined above.
Let $\hat{r} \coloneqq Y - \sum_{k \in S} \bar{\Delta}_k \hat{d}_k$ be the restricted regression residual. 
Suppose for all $k\in S^c$, for all $i=1,...,n$, $\lambda_n > | \frac{1}{2n}
\hat{r}^\tran \mathbf{1}_{(i:n)}|$ where $\mathbf{1}_{(i:n)}$ is 1 on
the coordinates of the 
$i$-th largest to the $n$-th largest entries of $X_k$ and 0
elsewhere.  Then the following two statements hold.
\begin{enumerate}
\item Let $\hat{d}_k = 0$ for $k \in S^c$.  Then
  \{$\hat{d}_k\}_{k=1,...,p}$ is an optimal solution to
  optimization~\eqref{opt:alternate_opt}. Furthermore, any solution to
  the optimization program \eqref{opt:alternate_opt} must be zero on
  $S^c$.
\item For all $k \in S^c$, the solution $\hat{c}_k$ to optimization~\eqref{opt:alternate_opt_concave} must be zero.
\end{enumerate}

@@

This result holds regardless of whether or not we impose the boundedness conditions in optimization~\eqref{opt:alternate_opt} and~\eqref{opt:alternate_opt_concave}.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.

Theorem~\ref{thm:deterministic} allows us to separately analyze the false negative
rates and false positive rates. To control false positives,
we analyze the condition on $\lambda_n$ for $k \in S^c$. To control
false negatives, we analyze the restricted regression.

The proof of Theorem~\ref{thm:deterministic} analyses the KKT
conditions of optimization~\eqref{opt:alternate_opt}.  This parallels
the now standard \emph{primal-dual witness}
technique~\cite{wainwright2009sharp}. However, we cannot derive analogous
\emph{mutual incoherence} conditions because the estimation is
nonparametric---even the low dimensional restricted regression has
$s(n-1)$ variables. The details of the proof are given in
Section~\ref{sec:deterministic_proof} of the Appendix.

@@

We let $S = \{ k = 1,...,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$.

@@

And therefore，for all $k \in S^c$, both the AC solution $\hat{f}_k$, from optimization~\eqref{opt:alternate_opt}, and the DC solution $\hat{g}_k$, from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}

@@

\begin{theorem} (Controlling false negatives) 
\label{thm:false_negative}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution to the restricted regression with $B$-boundedness constraint and let $\hat{g}_k$'s be any DC solution to the restricted regression with $B$-boundedness constraint. Let $\tilde{\sigma}$ denote $\max(\sigma, B)$.
